%*******************************************************
% Abstract in English
%*******************************************************
\pdfbookmark[0]{Abstract}{Abstract}


\begin{otherlanguage}{american}
	\chapter*{Abstract}
    % shorten first sentences so it's not the same as intro - immediately go into what this thesis will explore
	To extend the usability of the legacy pre-relational Adabas, the Event Replicator Target Adapter for Adabas on z/OS enables replication of Adabas files to modern systems, including relational databases. However, it faces performance challenges due to factors such as no support for parallel processing of replication events. The Kafka Connect framework was used to create an alternative to the Target Adapter to address its drawbacks, focusing mainly on the lack of parallelization. This research explores the question \textbf{"How does the performance of a Kafka Connect-based replication pipeline for Adabas on mainframe compare to the Adabas Event Replicator Target Adapter?"}. The two replication approaches were compared by evaluating their performance differences. Performance was be measured using JMX metrics and distributed tracing with Zipkin. Experiments included various configuration scenarios of the Kafka cluster against the single-process Target Adapter. The configuration scenarios were designed to explore how the degree of parallelization and distributedness affects the Kafka-based prototype's performance.
    % and simulated failures, such as worker node crashes, to observe recovery and rebalancing capabilities.
    The hypothesis was that the Kafka pipeline would outperform the Target Adapter if its distributed parallel processing capabilities were properly leveraged. If it would instead be run as a non-parallelized application like the Target Adapter, then the performance was expected to be equal or even worse, due to higher complexity and latency incurred by the multiple components.
    The experiment results demonstrated the positive effect of parallelization on the Kafka-based prototype's performance, with workload distribution also improving performance. The Kafka-based prototype out-performed the Target Adapter in both the single-threaded scenario and the highly parallelized scenarios.
    This research proves significant because it addresses a gap in the research concerning the modernization of pre-relational databases such as Adabas with modern approaches. Kafka is employed to facilitate parallel processing and a configurable integration with various systems. Such integration enables enhanced scalability, performance, and resilience in Adabas data replication, as well as improved integration of Adabas with modern and heterogeneous systems.
\end{otherlanguage}
